---
title: "AS3-2: 股市動態"
author: "第 10 組"
date: "`r Sys.time()`"
output:
  html_document:
    highlight: pygments
    theme: flatly
    css: style.css
---

```{r results='hide', message=FALSE, warning=FALSE, echo=FALSE}
# 這些程式碼設定網頁的格式，並安裝、載入一些基本的套件，請大家不要去改動<br>
rm(list=ls(all=T))
knitr::opts_chunk$set(comment = NA)
knitr::opts_knit$set(global.par = TRUE)
par(cex=0.8)
options(scipen=20, digits=4, width=90)
#
if(!require(pacman)) install.packages("pacman")
pacman::p_load(dplyr)
```
<hr>

**股票動態**

股票市場是買賣雙方交易公司股票的地方，也是個人和公司投資的最受歡迎的方式之一。現在估計世界股票市場規模達到數兆。紐約證券交易所位於紐約市，是世界上最大的股票市場。 紐約證券交易所約有2,800家上市公司。在這個問題上，我們將看看其中五家公司的每月股價：IB、通用電氣（GE）、寶潔、可口可樂和波音。此問題中使用的數據來自Infochimps。

使用`read.csv()`下載並讀取以下文件：

+ `data/IBMStock.csv`
+ `data/GEStock.csv`
+ `data/ProcterGambleStock.csv`
+ `data/CocaColaStock.csv`
+ `data/BoeingStock.csv` 

分別存入資料框`IBM`、`GE`、`ProcterGamble`、`CocaCola`和`Boeing`，
每個資料框都有兩個變量，描述如下：

+ `Date`：日期
+ `StockPrice`：給定月份公司的平均股價

在這個案例，我們將看看這些公司的股票動態如何隨著時間的推移而發生變化。

```{r}
IBM=read.csv('data/IBMStock.csv')
GE=read.csv('data/GEStock.csv')
PG=read.csv('data/ProcterGambleStock.csv')
CK=read.csv('data/CocaColaStock.csv')
BE=read.csv('data/BoeingStock.csv')
```


<br><hr>

###  Section-1 Summary Statistics

__§ 1.1__ Our five datasets all have the same number of observations. How many observations are there in each data set?
```{r}
head(IBM$Date)#Month/Day/Year
head(GE$Date)#Month/Day/Year
head(PG$Date)#Month/Day/Year
head(CK$Date)#Month/Day/Year
head(BE$Date)#Month/Day/Year

#480

#用head()查看資料
```


__§ 1.2__ What is the earliest year in our datasets?
```{r}
IBM$Date = as.Date(IBM$Date, format='%m/%d/%y')
GE$Date = as.Date(GE$Date, format='%m/%d/%y')
PG$Date = as.Date(PG$Date, format='%m/%d/%y')
CK$Date = as.Date(CK$Date, format='%m/%d/%y')
BE$Date = as.Date(BE$Date, format='%m/%d/%y')
head(as.Date(IBM$Date, format='%m/%d/%y'))
head(as.Date(GE$Date, format='%m/%d/%y'))
head(as.Date(PG$Date, format='%m/%d/%y'))
head(as.Date(CK$Date, format='%m/%d/%y'))
head(as.Date(BE$Date, format='%m/%d/%y'))

#1/1/1970

#用as.來轉換型態，並用head()來找尋最初五筆資料

```

__§ 1.3__ What is the latest year in our datasets?
```{r}
tail(as.Date(IBM$Date, format='%m/%d/%y'))
tail(as.Date(GE$Date, format='%m/%d/%y'))
tail(as.Date(PG$Date, format='%m/%d/%y'))
tail(as.Date(CK$Date, format='%m/%d/%y'))
tail(as.Date(BE$Date, format='%m/%d/%y'))

#12/1/2009

#用tail()來找尋最後五筆資料
```

__§ 1.4__ What is the mean stock price of IBM over this time period?
```{r}
mean(IBM$StockPrice)

#144.4

#用mean()來平均資料
```

__§ 1.5__ What is the minimum stock price of General Electric (GE) over this time period?
```{r}
min(GE$StockPrice)

#9.294

#用min()找資料最小值
```

__§ 1.6__ What is the maximum stock price of Coca-Cola over this time period?
```{r}
max(CK$StockPrice)

#146.6

#用max()找資料最大值
```

__§ 1.7__ What is the median stock price of Boeing over this time period?
```{r}
median(BE$StockPrice)

#44.88

#用median()找出中位數
```

__§ 1.8__ What is the standard deviation of the stock price of Procter & Gamble over this time period?
```{r}
sd(PG$StockPrice)

#331

#用sd找出標準差
```

<br><hr>

###  Section-2 Visualizing Stock Dynamics

__§ 2.1__ Around what year did Coca-Cola has its highest stock price in this time period? Around what year did Coca-Cola has its lowest stock price in this time period?
```{r}
plot(CK$Date, CK$StockPrice, type="l", main="COCACOLA")
abline(v=CK$Date[which.max(CK$StockPrice)], col="red")
abline(v=CK$Date[which.min(CK$StockPrice)], col="green")

#用plot()描繪出點狀圖，並在其中加入參數type="l"，將點連成線，成為折線圖
#用abline()加上輔助線，並用col=" "，以不同顏色標示
#ylim()為y的極限值，方便圖形判斷
#col=" "為指定顏色
#main=" "為圖形名稱
#xlab=" "為x軸名稱；ylab=" "為y軸名稱
#lwb= 為線條粗度

```


__§ 2.2__ In March of 2000, the technology bubble burst, and a stock market crash occurred. According to this plot, which company's stock dropped more?
```{r}
plot(CK$Date[361:373], CK$StockPrice[361:373], type="l", col="black", main="dropped more", xlab ="date", ylab="price", lwd=1, ylim = c(0, 210))
lines(BE$Date[361:373], BE$StockPrice[361:373], type="l", col="red", lwd=1)
lines(GE$Date[361:373], GE$StockPrice[361:373], type="l", col="green", lwd=1)
lines(IBM$Date[361:373], IBM$StockPrice[361:373], type="l", col="blue", lwd=1)
lines(PG$Date[361:373], PG$StockPrice[361:373], type="l", col="yellow", lwd=3)

#GE had droped more

#用plot()描繪出點狀圖，並在其中加入參數type="l"，將點連成線，成為折線圖。
#並指定資料區間，以找出想要的資料範圍，資料區間可由右上角的資料庫點開查看
#用lines()在原本的plot()點狀圖上加上其他直線
#ylim()為y的極限值，方便圖形判斷
#col=" "為指定顏色
#main=" "為圖形名稱
#xlab=" "為x軸名稱；ylab=" "為y軸名稱
#lwb= 為線條粗度

```

__§ 2.3__ (a) Around 1983, the stock for one of these companies (Coca-Cola or Procter and Gamble) was going up, while the other was going down. Which one was going up? 
```{r}
plot(CK$Date[145:180], CK$StockPrice[145:180], type="l", col="black", main="which one is loser", xlab = "date", ylab="price", ylim = c(0,150), lwd=1)
lines(PG$Date[145:180], PG$StockPrice[145:180], type = "l", col="yellow", lwd=3)

#CocaCola is going up

#用plot()描繪出點狀圖，並在其中加入參數type="l"，將點連成線，成為折線圖。
#並指定資料區間，以找出想要的資料範圍，資料區間可由右上角的資料庫點開查看
#用lines()在原本的plot()點狀圖上加上其他直線
#ylim()為y的極限值，方便圖形判斷
#col=" "為指定顏色
#main=" "為圖形名稱
#xlab=" "為x軸名稱；ylab=" "為y軸名稱
#lwb= 為線條粗度

```

(b) In the time period shown in the plot, which stock generally has lower values?
```{r}

#CocaCola is lower one

#用plot()描繪出點狀圖，並在其中加入參數type="l"，將點連成線，成為折線圖。
#並指定資料區間，以找出想要的資料範圍，資料區間可由右上角的資料庫點開查看
#用lines()在原本的plot()點狀圖上加上其他直線
#ylim()為y的極限值，方便圖形判斷
#col=" "為指定顏色
#main=" "為圖形名稱
#xlab=" "為x軸名稱；ylab=" "為y軸名稱
#lwb= 為線條粗度

```


<br><hr>

###  Section-3 Visualizing Stock Dynamics 1995-2005

__§ 3.1__ Which stock fell the most right after the technology bubble burst in March 2000?
```{r}
plot(CK$Date[301:432], CK$StockPrice[301:432], type="l", col="black", main="stock dynamics", xlab ="date", ylab="price", lwd=1, ylim = c(0, 210))
lines(BE$Date[301:432], BE$StockPrice[301:432], type="l", col="red", lwd=1)
lines(GE$Date[301:432], GE$StockPrice[301:432], type="l", col="green", lwd=1)
lines(IBM$Date[301:432], IBM$StockPrice[301:432], type="l", col="blue", lwd=1)
lines(PG$Date[301:432], PG$StockPrice[301:432], type="l", col="yellow", lwd=3)


#GE had the most fell in March 2000

#用plot()描繪出點狀圖，並在其中加入參數type="l"，將點連成線，成為折線圖。
#並指定資料區間，以找出想要的資料範圍，資料區間可由右上角的資料庫點開查看
#用lines()在原本的plot()點狀圖上加上其他直線
#ylim()為y的極限值，方便圖形判斷
#col=" "為指定顏色
#main=＂"為圖形名稱
#xlab=" "為x軸名稱；ylab=" "為y軸名稱
#lwb= 為線條粗度

```

__§ 3.2__ Which stock reaches the highest value in the time period 1995-2005?
```{r}

#IBM had reached the highest value in 1995-2005

#用plot()描繪出點狀圖，並在其中加入參數type="l"，將點連成線，成為折線圖。
#並指定資料區間，以找出想要的資料範圍，資料區間可由右上角的資料庫點開查看
#用lines()在原本的plot()點狀圖上加上其他直線
#ylim()為y的極限值，方便圖形判斷
#col=" "為指定顏色
#main=" "為圖形名稱
#xlab=" "為x軸名稱；ylab=" "為y軸名稱
#lwb= 為線條粗度


```

__§ 3.3__ In October of 1997, there was a global stock market crash that was caused by an economic crisis in Asia. Comparing September 1997 to November 1997, which companies saw a decreasing trend in their stock price? (Select all that apply.)
```{r}
plot(CK$Date[333:335], CK$StockPrice[333:335], type="l", col="black", main="decreasing trend", xlab ="date:1997", ylab="price", lwd=1, ylim = c(0, 210))
lines(BE$Date[333:335], BE$StockPrice[333:335], type="l", col="red", lwd=1)
lines(GE$Date[333:335], GE$StockPrice[333:335], type="l", col="green", lwd=1)
lines(IBM$Date[333:335], IBM$StockPrice[333:335], type="l", col="blue", lwd=1)
lines(PG$Date[333:335], PG$StockPrice[333:335], type="l", col="yellow", lwd=3)


#P&G had decreased in Sep. to Nov.

#用plot()描繪出點狀圖，並在其中加入參數type="l"，將點連成線，成為折線圖。
#並指定資料區間，以找出想要的資料範圍，資料區間可由右上角的資料庫點開查看
#用lines()在原本的plot()點狀圖上加上其他直線
#ylim()為y的極限值，方便圖形判斷
#col=" "為指定顏色
#main=" "為圖形名稱
#xlab=" "為x軸名稱；ylab=" "為y軸名稱
#lwb= 為線條粗度


```

__§ 3.4__ In the last two years of this time period (2004 and 2005) which stock seems to be performing the best, in terms of increasing stock price?
```{r}
plot(CK$Date[409:432], CK$StockPrice[409:432], type = "l", col="black", main = "stock trend", xlab = "date:2004-2005", ylab="price", ylim = c(0,210))
lines(BE$Date[409:432], BE$StockPrice[409:432], type="l", col="red")
lines(GE$Date[409:432], GE$StockPrice[409:432], type="l", col="green")
lines(IBM$Date[409:432], IBM$StockPrice[409:432], type="l", col="blue")
lines(PG$Date[409:432], PG$StockPrice[409:432], type="l", col="yellow", lwd=3)

#Boeing had best performance because it always go up

#用plot()描繪出點狀圖，並在其中加入參數type="l"，將點連成線，成為折線圖。
#並指定資料區間，以找出想要的資料範圍，資料區間可由右上角的資料庫點開查看
#用lines()在原本的plot()點狀圖上加上其他直線
#ylim()為y的極限值，方便圖形判斷
#col=" "為指定顏色
#main=" "為圖形名稱
#xlab=" "為x軸名稱；ylab=" "為y軸名稱
#lwb= 為線條粗度

```


<br><hr>


###  Section-4 Monthly Trends

__§ 4.1__ For IBM, compare the monthly averages to the overall average stock price. In which months has IBM historically had a higher stock price (on average)? Select all that apply.
```{r}
mean(IBM$StockPrice)#144.4
tapply(IBM$StockPrice,format(IBM$Date,'%m'),mean)

#January、February、March、April、May. There have five months stock price higher than 144.4

#用mean()找出月股價平均值
#用tapply()將IBM的每月股價進行平均，再進行比較

```

__§ 4.2__ General Electric and Coca-Cola both have their highest average stock price in the same month. Which month is this?
```{r}
which.max(tapply(GE$StockPrice,format(GE$Date,'%m'),mean))#
which.max(tapply(CK$StockPrice,format(CK$Date,'%m'),mean))#

#Both of April average stock price are highest

#用tapply()將GE、CK的每月股價進行平均
#用which.max()找出最高平均股價的月份

```

__§ 4.3__ For the months of December and January, every company's average stock is higher in one month and lower in the other. In which month are the stock prices lower?
```{r}
tapply(IBM$StockPrice,format(IBM$Date,'%m'),mean)
tapply(GE$StockPrice,format(GE$Date,'%m'),mean)
tapply(CK$StockPrice,format(CK$Date,'%m'),mean)
tapply(PG$StockPrice,format(PG$Date,'%m'),mean)
tapply(BE$StockPrice,format(BE$Date,'%m'),mean)

#All of the December average stock price are lower than January in five companys

#用tapply()將IBM、GE、CK、PG、BE的每月股價進行平均，再進行比較

```

<br><br><br>

